\LARGE \bf Informed sampling by Markov Chain Monte Carlo methods

### 2017-02-22 paper.tex
added alg

\section{MCMC-based Informed Sampling}
In this section we describe our approach to efficiently produce samples in the informed set~\Cinf given a specific cost~$c_{\text{best}}$.
We start  by randomly sampling a state~$x_\text{rand} \in \calX$ and use it
to produces a sample~$x_\text{boundary} \in \partial\Cinf$ which lies (approximately) on the boundary of the informed set (Sec.~\ref{subsec:grad}).
This is done by applying Newton–Raphson method~\cite{RT06}.
We then use~$x_\text{boundary}$ to initialize an MCMC algorithm which produces a series of samples within \Cinf (Sec.~\ref{subsec:mcmc}). Specifically, we chose to use HMC due to its favourable properties (see Sec.~\ref{sec:intro}).

\subsection{Sampling the boundary of \Cinf via Newton–Raphson}
\label{subsec:grad}
In theory, MCMC methods converge to the desired distribution regardless of the initial sample used to seed the chain.
However, in practice it is common to discard an initial set of samples (a process referred to as ``burn-in'') to avoid starting biases~\cite{ADDJ03}.

In our setting, we are only interested in points in~\Cinf, thus we can start the Markov Chain in~\Cinf and avoid this burn-in stage. 
To do so we apply  Newton–Raphson's method to obtain a sample on $\partial \Cinf$. 
However, our cost function is not necessarily continuous (see Sec.~\ref{sec:mtdi}), thus if no sample was produced on $\partial \Cinf$ after a predefined number of steps, we restart the algorithm using a different random state~$x_\text{rand}$.

@@

The term 
$E_{\text{informed}} \left( t,\Cbest \right) 
= \beta \cdot \text{sigm}\left(t_0 - t\right)
$
penalizes states outside the informed set~$\Cinf$ using a tight sigmoid function~$\text{sigm}$ scaled by some (large\footnote{The constant~$\beta$ gives points out of~\Cinf a large energy value which in turn implies a low probability of being sampled. In theory, we would like to have~$\beta \approx \infty$. However, choosing~$\beta$ to be to large may cause numerical instabilities.}) constant~$\beta$.
The sigmoid can be seen as a differentiable function smoothing the boundary of~$\Cinf$.

\os{add visualization}

>>>>

\section{MCMC-based Informed Sampling}
In this section we describe our approach to efficiently produce samples in the informed set~\Cinf given a specific cost~$c_{\text{best}}$.
We start  by randomly sampling a state~$x_\text{rand} \in \calX$ and use it
to produces a sample~$x_\text{boundary} \in \partial\Cinf$ which lies (approximately) on the boundary of the informed set (Sec.~\ref{subsec:grad}).
This is done by applying Newton-Raphson method~\cite{RT06}.
We then use~$x_\text{boundary}$ to initialize an MCMC algorithm which produces a series of samples within \Cinf (Sec.~\ref{subsec:mcmc}). Specifically, we chose to use HMC due to its favourable properties (see Sec.~\ref{sec:intro}).

\begin{algorithm}[t]
	\begin{algorithmic}[1]
		\STATE $x_{\text{rand}} \leftarrow \texttt{sample\_uniform(\calX)}$
		\STATE $x_{\text{boundary}} \leftarrow
		 \texttt{newton\_raphson}(x_{\text{rand}}, c_{\text{best}})$
		\STATE $x_{0} \leftarrow x_{\text{boundary}}$
		\FOR {$i = 1$ to $N$}
			\STATE $x_{i} \leftarrow \texttt{MCMC} (x_{i-1}, c_{\text{best}})$
			\RETURN $x_{i}$
		\ENDFOR

   	\end{algorithmic}
	\caption{MCMC Informed Sampling $(c_{\text{best}})$}
	\label{alg:mcmc_informed_sampling}	
\end{algorithm}

\subsection{Sampling the boundary of \Cinf via Newton-Raphson}
\label{subsec:grad}
In theory, MCMC methods converge to the desired distribution regardless of the initial sample used to seed the chain.
However, in practice it is common to discard an initial set of samples (a process referred to as ``burn-in'') to avoid starting biases~\cite{ADDJ03}.

In our setting, we are only interested in points in~\Cinf, thus we can start the Markov Chain in~\Cinf and avoid this burn-in stage. 
To do so we apply  Newton-Raphson's method to obtain a sample on $\partial \Cinf$. 
However, our cost function is not necessarily continuous (see Sec.~\ref{sec:mtdi}), thus if no sample was produced on $\partial \Cinf$ after a predefined number of steps, we restart the algorithm using a different random state~$x_\text{rand}$.

@@

The term 
$E_{\text{informed}} \left( t,\Cbest \right) 
= \beta \cdot \text{sigm}\left(t_0 - t\right)
$
penalizes states outside the informed set~$\Cinf$ using a tight sigmoid function~$\text{sigm}$ scaled by some (large\footnote{The constant~$\beta$ gives points out of~\Cinf a large energy value which in turn implies a low probability of being sampled. In theory, we would like to have~$\beta \approx \infty$. However, choosing~$\beta$ to be to large may cause numerical instabilities.}) constant~$\beta$.
The sigmoid can be seen as a differentiable function smoothing the boundary of~$\Cinf$.
See Fig.~\ref{fig:energy} for a visualization.

\begin{figure}[tb]
  \centering
  	\includegraphics[height = 4.5cm ]{fig/energy.pdf}
  \caption{
    \captionstyle
  	Energy function used by HMC to obtain samples~$x \in \Cinf$.
  	}
   	\label{fig:energy}
\end{figure}
